### 2021-09-14 notes.tex
Add global notes with kaobook style

\documentclass[
	fontsize=11pt, % Base font size
	twoside=false, % Use different layouts for even and odd pages (in particular, if twoside=true, the margin column will be always on the outside)
	%open=any, % If twoside=true, uncomment this to force new chapters to start on any page, not only on right (odd) pages
	%chapterprefix=true, % Uncomment to use the word "Chapter" before chapter numbers everywhere they appear
	%chapterentrydots=true, % Uncomment to output dots from the chapter name to the page number in the table of contents
	numbers=noenddot, % Comment to output dots after chapter numbers; the most common values for this option are: enddot, noenddot and auto (see the KOMAScript documentation for an in-depth explanation)
	%draft=true, % If uncommented, rulers will be added in the header and footer
	%overfullrule=true, % If uncommented, overly long lines will be marked by a black box; useful for correcting spacing problems
]{kaohandt}

% Choose the language
\usepackage[english]{babel} % Load characters and hyphenation
\usepackage[english=british]{csquotes}	% English quotes

% Load the bibliography package
\usepackage{kaobiblio}
\addbibresource{bibliography.bib} % Bibliography file

% Load the package for hyperreferences
\usepackage{kaorefs}

% Set the paths where to look for images
\usepackage{subcaption}
\graphicspath{{figures/}}

\begin{document}

\title{Notes}
\author[FM]{Antonio Coín Castro}
\date{\today}
\maketitle

\defbibnote{bibnote}{Here are the references in citation order.\par\bigskip} % Prepend this text to the bibliography
\renewcommand*{\bibfont}{\small}
\printbibliography[title=Bibliography]% , prenote=bibnote]

\end{document}