### 2022-06-24 slides/seio.tex

%
% Copyright (c) 2021 Antonio Coín Castro
%
% This work is licensed under a
% Creative Commons Attribution-ShareAlike 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by-sa/4.0/>.
%

\RequirePackage{fix-cm}
\documentclass[10pt, english, professionalfonts]{beamer}

\usepackage{pifont}  % xmark, cmark
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

% OPCIONES DE BEAMER

\definecolor{Maroon}{cmyk}{0, 0.87, 0.88, 0.1}
\definecolor{teal}{rgb}{0.0, 0.45, 0.45}

\usetheme[block=fill, subsectionpage=progressbar, titleformat section=smallcaps]{metropolis}
\setbeamertemplate{frametitle continuation}[roman]
\setbeamertemplate{section in toc}[sections numbered]
%\setbeamertemplate{subsection in toc}[subsections unnumbered]
%\setsansfont[BviejoFont={Fira Sans SemiBold}]{Fira Sans Book}  % Increase font weigth
\widowpenalties 1 10000
\raggedbottom

% COLORES
\setbeamercolor{palette primary}{bg=teal}
\setbeamercolor{progress bar}{use=Maroon, fg=Maroon}

% PAQUETES

\usepackage[utf8]{inputenc}
\usepackage[absolute,overlay]{textpos}
\usepackage[english]{babel}
\usepackage{microtype}
\usepackage{epigraph}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{amssymb, amsmath, amsthm, amsfonts, amscd}
\usepackage{listings}

\usepackage[backend=biber,style=numeric-comp,sorting=none]{biblatex}
\addbibresource{bibliography.bib}
\setbeamertemplate{bibliography item}[article]

% FONTS
%\usefonttheme{professionalfonts}
%\usepackage{mathpazo}
%\usepackage{eulervm}

\definecolor{backg}{HTML}{F2F2F2} % Fondo
\definecolor{comments}{HTML}{a8a8a8} % Comentarios
\definecolor{keywords}{HTML}{08388c} % Palabras clave
\definecolor{strings}{HTML}{0489B1}  % Strings

\lstset{
language=scala,
basicstyle=\footnotesize\ttfamily,
breaklines=true,
keywordstyle=\color{keywords},
commentstyle=\color{comments},
stringstyle=\color{strings},
xleftmargin=.5cm,
tabsize=2,
% Acentos, ñ, ¿, ¡ (tex.stackexchange.com/questions/24528)
extendedchars=true
}

% COMANDOS PERSONALIZADOS

\renewcommand{\baselinestretch}{1}
\definecolor{mLightBrown}{HTML}{f97e0b}
\newcommand\maroon[1]{\color{mLightBrown}#1\color{mDarkTeal}}
\let\lmin\wedge
\let\lmax\vee
\newtheorem{prop}{Proposición}
\newtheorem{teorema}{Teorema}
\newtheorem{defi}{Definición}
\newcommand\ddfrac[2]{\frac{\displaystyle #1}{\displaystyle #2}}  % Fracción grande

\renewcommand{\epsilon}{\varepsilon}

\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Hcal}{\ensuremath\mathcal{H}}

%% Scalar product
\newcommand\dotprod[2]{\left\langle #1, #2 \right\rangle}

% PLOTS

%\usepackage{pgfplots}
%\DeclareUnicodeCharacter{2212}{−}
%\usepgfplotslibrary{groupplots,dateplot}
%\usetikzlibrary{patterns,shapes.arrows}
%\pgfplotsset{compat=newest}

% TÍTULO

\title{Bayesian RKHS-based methods in functional regression}
\providecommand{\subtitle}[1]{}
\subtitle{Theoretical framework and experiments}
%\date{SEIO Conference, Granada\\ \today\\}
\date{\today\\ \\}
\author{José R. Berrendero \\ Antonio Coín  \\ Antonio Cuevas\\}
\institute{Universidad Autónoma de Madrid \\ \textit{Departamento de Matemáticas}}

\titlegraphic{
  \begin{textblock*}{2cm}(9.8cm, 6.7cm)
    \includegraphics[width=2cm]{img/logo-uam}
  \end{textblock*}
}
% DOCUMENTO

\begin{document}
\maketitle

\begin{frame}{Table of contents}
   \tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}{FDA: linear regression}
  \textbf{First problem}: \maroon{linear } regression with functional data.
  \vspace{1em}

\begin{figure}
    \includegraphics[width=\textwidth]{img/data_lin}
    \caption{Linear regression data set simulated from a fractional Brownian motion.}
  \end{figure}
  \vspace{-1em}
\end{frame}
\begin{frame}{FDA: logistic regression}
  \textbf{Second problem}: \maroon{logistic } regression with functional data.
\vspace{1em}

\begin{figure}
    \includegraphics[width=\textwidth]{img/data_log}
    \caption{Logistic regression data set simulated from a fractional Brownian motion.}
  \end{figure}
  \vspace{-1em}
\end{frame}

\begin{frame}{The usual \(\bm{L^2}\)-models}
  \(\bullet\) \(Y\) real variable (continuous or dichotomous).

  \(\bullet\) \(X(t)\) second order stochastic process with trajectories in \(L^2[0, 1]\).

  \(\bullet\) \(\beta \in L^2[0,1]\) \maroon{functional parameter}.

  \vspace{1em}

\begin{block}{Linear \(\bm{L^2}\)-model}
  \[
    Y = \alpha_0 + \maroon{\langle X, \beta \rangle_2} + \varepsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \varepsilon.
  \]
\end{block}
\setbeamerfont{block body}{size=\small}
\begin{block}{Logistic \(\bm{L^2}\)-model}
    \[
    \mathbb P (Y=1\mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X, \beta\rangle_2}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \int_0^1 X(t)\beta(t)\, dt\}}.
  \]
\end{block}
  \(\alpha_0\in\mathbb R\), \ \(\varepsilon \sim \mathcal N(0, \sigma^2), \ X \bot \varepsilon\).
  \setbeamerfont{block body}{size=\normalsize}
\end{frame}

\begin{frame}{Shortcomings of the \(\bm{L^2}\)-models}
  \begin{itemize}
    \item \(L^2[0, 1]\) is a extremely wide space, which also contains very ill-behaved functions.
    \item The usual least-squares theory cannot be applied directly; some regularization or dimensionality reduction techniques are required.
    \item They do not include simple models based on linear combinations of the marginal of the process as particular cases, such as
    \[
      Y = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \varepsilon.
    \]
    \item In the logistic case, the maximum likelihood estimator does not exist with probability one under fairly general conditions.
  \end{itemize}
\end{frame}

\section{A RKHS model for functional regression}

\begin{frame}{Alternative model: RKHS}

We will exploit the theory of \textit{reproducing kernel Hilbert spaces} (RKHS's).

\vspace{1em}
\begin{alertblock}{Proposal}
  Change the habitat of the functional parameter \(\beta(\cdot)\).
\end{alertblock}

\vspace{1em}
  Instead of assuming \(\beta \in L^2[0, 1]\), we consider \maroon{\(\beta \in \mathcal H(K)\)}, the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\) (i.e. \(\E [X(t)]=0\) for all \(t\)). Moreover, we substitute \(\dotprod{X}{\beta}_2\) for \maroon{\(\dotprod{X}{\beta}_K\)}.
\end{frame}

\begin{frame}{RKHS reminder}
  \begin{definition}
    Let the covariance function \(K\) be continuous, and define the preliminary space
    \[
    \Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot),  p \in \N,  a_i \in \R,  t_i \in [0, 1] \right\},
    \]
    which is endowed with the scalar product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). Then, \(\Hcal(K)\) is the completion of \(\Hcal_0(K)\) under this scalar product.
  \end{definition}

  \vspace{1em}

  Note that \(\Hcal(K)\subset L^2[0, 1]\) as subsets, but the elements of \(\Hcal(K)\) are simpler and more manageable in general. Plus, the latter is a space of \maroon{functions}, and not of equivalence classes.
\end{frame}

\begin{frame}{RKHS reminder (cont.)}
  There are other equivalent definitions of \(\Hcal(K)\), mainly based on the \textit{covariance operator}
\[
\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f\in L^2[0, 1].
\]

\begin{definition}
\begin{enumerate}
  \item  \(\Hcal(K) = \mathcal K^{1/2}(L^2[0, 1])\), with \(\dotprod{f}{g}_K = \dotprod{\mathcal K^{-1/2}(f)}{\mathcal K^{-1/2}(g)}_2\).
  \item \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}_2\). Here, \(\lambda_j\) and \(\phi_j\) are the eigenvalues and (orthonormal) eigenfunctions of \(\mathcal K\).
\end{enumerate}
\end{definition}

The second definition emphasizes the fact that functions in \(\Hcal(K)\) are ``simple'', in the sense that their components in an orthonormal basis need to vanish quickly.

\end{frame}

\begin{frame}{A small subtlety}

  Generally speaking, the concrete realizations \(x=X(\omega)\) \maroon{do not belong to \(\Hcal(K)\) with probability \(1\)}, so the expression \(\dotprod{x}{\beta}_K\) lacks meaning. However, we can make use of the relationship between \(\Hcal(K)\) and \(\mathcal L(X)\), the linear span of the process \(X\) within \(L^2(\Omega)\).

  \vspace{1em}
  \begin{block}{Loève's isometry}
    It is the completion of the correspondence
    \[
\sum_{j=1}^p a_j X(t_j) \longleftrightarrow \sum_{j=1}^p a_j K(t_j, \cdot).
    \]
    Formally, \(\Psi^{-1}_X(U)(t) := \E[U X(t)]\) for \(U \in \mathcal L(X)\).
  \end{block}
  With a slight abuse of notation, we will write \maroon{\(\dotprod{x}{\beta}_K \equiv [\Psi_X(\beta)](\omega) := \Psi_x(\beta)\)}\ for \(x=X(\omega)\). This correspondence will allow us to still call our models ``linear''.
\end{frame}

\begin{frame}{A new functional parameter...}
  \begin{itemize}
    \item The initial idea was to suppose \(\beta \in \Hcal(K)\).
    \item To further simplify the model, we will assume \(\beta \in \Hcal_0(K)\), which is dense in \(\Hcal(K)\) by construction. That is,
    \[
      \beta(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot), \quad \beta_j \in \R, \ t_j \in [0, 1], \ p \in \N.
    \]
    \item When evaluating \(\dotprod{x}{\beta}_K\) for a trajectory \(x=X(\omega)\), we resort to Loève's isometry:
    \[
      \dotprod{x}{\beta}_K = \Psi_x(\beta) = \sum_{j=1}^p \beta_j X(t_j).
    \]
  \end{itemize}

\end{frame}

\begin{frame}{...and two new models}
  Given a data set \(\mathcal D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), for \(\beta \in \Hcal_0(K)\) we consider:

  \vspace{1em}

  \begin{block}{Linear RKHS model (Berrendero et al., 2019)}
  \[
    Y_i = \alpha_0 + \maroon{\langle X_i, \beta \rangle_K} + \varepsilon = \alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j) + \varepsilon.
  \]
\end{block}
\vspace{1em}

\setbeamerfont{block body}{size=\small}
\begin{block}{Logistic RKHS model (Berrendero et al., 2018)}
    \[
    \mathbb P (Y_i=1| X_i) = \frac{1}{1 + \exp\{-\alpha_0 - \maroon{\langle X_i, \beta\rangle_K}\}} = \frac{1}{1 + \exp\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\}}.
  \]
\end{block}
\setbeamerfont{block body}{size=\normalsize}

\end{frame}

\begin{frame}{Advantages of RKHS models}
  \begin{itemize}
    \item They are finite-dimensional approximations with a functional interpretaion, which simplifies their understanding and handling.
    \item Under mild conditions, the general RKHS models include the \(L^2\)-models as particular cases.
    \item The problem of non-existence of the MLE is mitigated in the logistic model.
    \item They are particularly suited as a basis for variable selection methods.
  \end{itemize}
\end{frame}

\section{A bayesian approach to parameter estimation}

\begin{frame}{Parameter estimation overview}

  We have a parameter vector
  \[
  \theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2)' \equiv (p, b, \tau, \alpha_0, \sigma^2)',
  \]
  on which we impose a prior distribution \(\pi(\theta)\). In practice, we are establishing a prior distribution on the functional space \(\Hcal(K)\) via a density argument.

  \vspace{1em}

  Afterwards we apply \textit{Bayes'} theorem in the usual way to obtain the posterior distribution of the parameters given the data.

  \begin{block}{Bayes's theorem}
      \[
      \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
      \]
  \end{block}

\end{frame}

\begin{frame}{Prior distributions}
  Same prior scenario for both linear and logistic models.
  \vspace{.5em}
  \begin{block}{Prior for the parameters involved}
    \vspace{-1em}
  \begin{align*}
    p &\sim \text{Categorical}(p_{\text{max}}),\\
  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
  b\mid \tau, \sigma^2 & \sim \mathcal N_p\big(b_0, g\sigma^2{\underbrace{\left[\mathcal X_\tau' \mathcal X_\tau + \eta I\right]}_{G_\tau}}^{-1}\big),
\end{align*}

\vspace{-1em}
where \(I\) is the identity matrix, \(\mathcal X_\tau = (X_i(t_j))_{i,j}\) is the data matrix, and \(b_0\in \R^p, \ g \in \R, \ p_{\text{max}}\in \N\) and \(\eta \in \R^+\) are hyperparameters.
\end{block}

\metroset{block=transparent}

\vspace{1em}
\begin{alertblock}{Hyperparameter selection}
  \vspace{0.1em}
  For \(b_0\) we use a computational approximation of the MLE; the rest of hyperparameters are set via \textit{cross-validation}.
\end{alertblock}
\metroset{block=fill}
\end{frame}

\begin{frame}{Posterior distributions}
  \setbeamerfont{block body}{size=\small}
  \begin{block}{Linear log-posterior}
    \vspace{-1em}
    \begin{align*}
  \log \pi(\theta\mid \mathcal D_n) &\propto \frac{1}{2}\log |G_\tau| - (p+n)\log \sigma\\
  &-\frac{1}{2\sigma^2} \left(\|\boldsymbol{Y}-\alpha_0\boldsymbol{1} - \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right),
\end{align*}
where \(\bm Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
  \end{block}
  \vspace{1em}
  \begin{block}{Logistic log-posterior}
      \vspace{-1em}
    \begin{align*}
  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
  \quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
\end{align*}
Remember that \(\Psi_{X_i}(\alpha) = \sum_j \beta_j X_i(t_j)\).
  \end{block}
  \setbeamerfont{block body}{size=\normalsize}
  They are approximated through \textit{Markov chain Monte Carlo} (MCMC).
\end{frame}

\begin{frame}{Posterior approximation}
 We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).

  \vspace{1em}

  \begin{figure}
    \includegraphics[width=.7\textwidth]{img/posterior}
    \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
  \end{figure}
\end{frame}

\begin{frame}{Predictions}
  \textbf{Summarize-then-predict}. A summary of the posterior distribution \(\theta| \mathcal D_n\) is obtained via a point-estimate statistic (mean, median, mode), and then predictions are made according to the assumed model, e.g.:
  \[
  \hat Y_i =\hat \alpha_0 + \sum_{j=1}^p \hat \beta_j X_i(\hat t_j), \quad i=1,\dots, n.
  \]

  \textbf{Predict-then-summarize}. We first construct predictions on each step of the MCMC algorithm following our model, i.e.:
  \[
    Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}, \quad m=1,\dots,M, \quad i=1,\dots, n.
  \]
  Then, the mean of all such intermediate predictions (\textit{posterior predictive distribution}) is used as a proxy for the response variable:
  \[
    \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots, n.
  \]

\end{frame}

\begin{frame}{Predictions (cont.)}
    \vspace{1em}
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/ppc_lin_standalone}
    \caption{Example of \textit{posterior predictive distribution} in linear regression.}
  \end{figure}
\end{frame}

\begin{frame}{Predictions (cont.)}
    \textbf{Variable selection}. We select \(p\) variables using only a point estimator of \(\tau|\mathcal D_n\) (so-called \textit{impact points}), and then we can apply any multiple regression algorithm to the reduced data set.
    \vspace{1em}

    \begin{figure}
      \includegraphics[width=\textwidth]{img/tau_posterior}
      \caption{Approximate posterior distribution for the impact points.}
    \end{figure}

    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
\end{frame}

\begin{frame}{Bayesian model checking}
  \begin{itemize}
    \item It is possible to incorporate existing information through the prior distributions.
    \item Analysis of the trace and posterior distribution of the parameters, for example through credible intervals.
    \item \textit{Bayesian p-values}: \(\tilde p=P(T(Y^*)\leq T(Y)\mid Y)\) for any choice of \(T\) (min, max, mean, median, mode, etc.). They are computed by simply measuring the proportion of the approximate posterior samples that fall below the real value of the statistic, and are expected to be around \(0.5\) when the fit is good.
    \item Descriptive and visual analysis of the posterior predictive distribution, both with training and validation data.
  \end{itemize}

\end{frame}

\section{Experiments}

\begin{frame}{Methodology}
  \begin{itemize}
    \item Experiments on simulated data, with \(300\) samples of different Gaussian processes on a regular grid of \(100\) points. Also experiments on real data sets.
    \item We perform 10 random train/test partitions (67\%/33\%) to account for the stochasticity of the model, along with \(5\)-fold cross-validation to select the best hyperparameters on the training set.
    \item For practical and computational reasons, \textbf{we allow the value of \(\bm{p}\) to be fixed beforehand}, treating it as an hyperparameter (equivalent to a degenerate prior distribution).
    \item Several reference methods are considered for comparison, along with variable selection techniques, functional or otherwise.

  \end{itemize}
\end{frame}

\begin{frame}{Linear regression: simulated RKHS data}
  \vspace{1em}
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/reg_emcee_rkhs}
    \caption{Mean MSE of predictors for simulated GP data that obeys an underlying RKHS model (left is better).}
  \end{figure}
\end{frame}

\begin{frame}{Linear regression: simulated \(\bm{L^2}\) data}
  \vspace{1em}
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/reg_emcee_l2}
    \caption{Mean MSE of predictors for simulated GP data that obeys an underlying \(L^2\)-model (left is better).}
  \end{figure}
\end{frame}

\begin{frame}{Linear regression: real data}
    \vspace{1em}
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/reg_emcee_real}
    \caption{Mean MSE of predictors for real data sets (left is better).}
  \end{figure}
\end{frame}

\begin{frame}{Logistic regression: simulated RKHS data}
    \vspace{1em}
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/clf_emcee_rkhs}
    \caption{Mean accuracy of classifiers for simulated GP data that obeys an underlying logistic RKHS model (right is better).}
  \end{figure}
\end{frame}

\begin{frame}{Logistic regression: simulated \(\bm{L^2}\) data}
    \vspace{1em}
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/clf_emcee_l2}
    \caption{Mean accuracy of classifiers for simulated GP data that obeys an underlying logistic \(L^2\)-model (right is better).}
  \end{figure}
\end{frame}

\begin{frame}{Logistic regression: real data}
    \vspace{1em}
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/clf_emcee_real}
    \caption{Mean accuracy of classifiers for real data sets (right is better).}
  \end{figure}
\end{frame}

\section{Conclusions}

\begin{frame}{Conclusions and future work}

  \metroset{block=transparent}
  \begin{alertblock}{Our proposal}
    \begin{itemize}
      \item Simple and very interpretable functional regression model, with a solid mathematical background thanks to the theory of RKHS's.
      \item Computationally feasible Bayesian approach to parameter estimation within the model.
      \item Competitive with usual alternatives in both simulated and real data sets.
    \end{itemize}
  \end{alertblock}

  \begin{exampleblock}{The road ahead}
    \begin{itemize}
      \item Experiment with different prior distributions.
      \item Derive theoretical properties of the predictors.
      \item Package the Python algorithm and make it compatible with standard libraries (e.g. \textit{sklearn}, \textit{scikit-fda}).
    \end{itemize}
  \end{exampleblock}

\end{frame}

\begin{frame}[allowframebreaks]
    \frametitle{References}
    \nocite{*}
    \printbibliography[heading=none]
\end{frame}

\begin{frame}[standout]
  Thank you!
\end{frame}

\end{document}